### 2021-10-20 paper/pldi22/formal.tex
Outline for formalization of PE

\subsection{Abstract Syntax}

\[
  \begin{array}{lr@{~}c@{~}c@{~}c@{~}l}
    \text{Programs} & \progset &\ni& \prog &::=& \langle \func^*, \inst \rangle\\
    \text{Functions} & \funcset &\ni& \func &::=&
    \kwdef \; x \; \kwrl x^* \kwrr \; \kwcl [\lab: \inst]^* \kwcr\\
    \text{Variables} & \varset &\ni& x\\
    \text{Labels} & \labset &\ni& \lab\\
    \text{Instructions} & \instset &\ni& \inst &::=&
    x = \pval \mid
    x = \op \kwrl x^* \kwrr \mid
    x = x \kwrl x^* \kwrr\\
    &&&&\mid&
    x = \kwcl \kwcr \mid
    x = x \kwsl x \kwsr \mid
    x \kwsl x \kwsr = x\\
    &&&&\mid&
    \kwif \; x \; \lab \; \lab \mid
    \kwret \; x\\
  \end{array}
\]

A program $\prog \in \progset$ is a pair of functions and a top-level
instruction.  A function $\func \in \funcset$ is defined with its name,
parameters, and body instructions with labels $\labset$.  An instruction $\inst
\in \instset$ is
a primitive value assignment $x = \pval$,
an operation between values $x = \op \kwrl x^* \kwrr$,
a function call $x = x \kwrl x^* \kwrr$,
a record allocation $x = \kwcl \kwcr$,
a field access $x = x \kwsl x \kwsr$,
a field assignment $x \kwsl x \kwsr = x$,
a conditional branch $\kwif \; x \; \lab \; \lab$,
or a return instruction $\kwret \; x$.

\subsection{States}

\[
  \begin{array}{lr@{~}c@{~}l@{~}c@{~}l}
    \text{States} & \st &\in& \stset &=&
    \heapset \times \ctxtset\\
    \text{Heaps} & \heap &\in& \heapset &=&
    \addrset \finmap \recset\\
    \text{Contexts} & \ctxt &\in& \ctxtset &=&
    \labset \times \envset \times \ctxtset^?\\
    \text{Environments} & \env &\in& \envset &=&
    \varset \finmap \valset\\
    \text{Values} & \val &\in& \valset &=&
    \pvalset \uplus \addrset \uplus \treeset \uplus \funcset\\
    \text{Primive Values} & \pval &\in& \pvalset &=&
    \strset \uplus \cdots\\
    \text{JavaScript ASTs} & \tree &\in& \treeset\\
    \text{Records} & \rec &\in& \recset &=&
    \strset \finmap \valset\\
  \end{array}
\]

States $\stset$ consists of heaps $\heapset$ and contexts $\ctxtset$.  A heap
$\heap \in \heapset$ is a finite mapping from addresses to records.  Each record
allocation instruction $x = \kwcl \kwcr$ creates a unique address $\addr \in
\addrset$ different from existing addresses.  A context $\ctxt \in \ctxtset$ is
a triple of 1) a program point, 2) an environment, and 3) an optional calling
context.  An environment $\env \in \envset$ is a finite mapping from variables
to values. A value $\val \in \valset$ is a primitive value $\pval \in \pvalset$
(e.g., a string value $\str \in \strset$), an address $\addr \in \addrset$, an
JavaScript AST $\tree \in \treeset$, or a function $f \in \funcset$.  A record
$\rec \in \recset$ is a finite mappings from strings to values.

>>>>

We first define $\ires$ with its abstract syntax, states, and concrete
semantics.  Then, we define syntactic views as JavaScript ASTs with abstract
nodes. Finally, we define the partial evaluation as a transformation of $\ires$
to specialize $\ires$ functions.  To prove the semantics preservation of the
partial evaluation, we introduce an observational semantics of $\ires$ and prove
that each specialized $\ires$ functions has equivalent observational semantics
of the corresponding original one.

\subsection{$\ires$: Intermediate Representations for ECMAScript}

First, we define abstract syntax, states, and concrete semantics of $\ires$.

\subsubsection{Abstract Syntax}

\[
  \begin{array}{lr@{~}c@{~}c@{~}c@{~}l}
    \text{Functions} & \funcset &\ni& \func &::=&
    \kwdef \; x \; \kwrl x^* \kwrr \; \kwcl [\lab: \inst]^* \kwcr\\
    \text{Variables} & \varset &\ni& x\\
    \text{Labels} & \labset &\ni& \lab\\
    \text{Instructions} & \instset &\ni& \inst &::=&
    x = \pval \mid
    x = \op \kwrl x^* \kwrr \mid
    x = x \kwrl x^* \kwrr\\
    &&&&\mid&
    x = \kwcl \kwcr \mid
    x = x \kwsl x \kwsr \mid
    x \kwsl x \kwsr = x\\
    &&&&\mid&
    \kwif \; x \; \lab \; \lab \mid
    \kwret \; x\\
  \end{array}
\]

A function $\func \in \funcset$ is defined with its name, parameters, and body
instructions with labels $\labset$.  An instruction $\inst \in \instset$ is
a primitive value assignment $x = \pval$,
an operation between values $x = \op \kwrl x^* \kwrr$,
a function call $x = x \kwrl x^* \kwrr$,
a record allocation $x = \kwcl \kwcr$,
a field access $x = x \kwsl x \kwsr$,
a field assignment $x \kwsl x \kwsr = x$,
a conditional branch $\kwif \; x \; \lab \; \lab$,
or a return instruction $\kwret \; x$.

\subsubsection{States}

\[
  \begin{array}{lr@{~}c@{~}l@{~}c@{~}l}
    \text{States} & \st &\in& \stset &=&
    \heapset \times \envset \times \ctxtset\\
    \text{Heaps} & \heap &\in& \heapset &=&
    \addrset \finmap \recset\\
    \text{Contexts} & \ctxt &\in& \ctxtset &=&
    \labset \times \envset \times \ctxtset^?\\
    \text{Environments} & \env &\in& \envset &=&
    \varset \finmap \valset\\
    \text{Values} & \val &\in& \valset &=&
    \pvalset \uplus \addrset \uplus \treeset \uplus \funcset\\
    \text{Primive Values} & \pval &\in& \pvalset &=&
    \strset \uplus \cdots\\
    \text{JavaScript ASTs} & \tree &\in& \treeset\\
    \text{Records} & \rec &\in& \recset &=&
    \strset \finmap \valset\\
  \end{array}
\]

States $\stset$ consist of heaps $\heapset$, global environments $\envset$, and
contexts $\ctxtset$.  A heap $\heap \in \heapset$ is a finite mapping from
addresses to records.  Each record allocation instruction $x = \kwcl \kwcr$
creates a unique address $\addr \in \addrset$ different from existing addresses.
A context $\ctxt \in \ctxtset$ is a triple of a label, a local environment, and
an optional calling context.  An environment $\env \in \envset$ is a finite
mapping from variables to values. A value $\val \in \valset$ is a primitive
value $\pval \in \pvalset$ (e.g., a string value $\str \in \strset$), an address
$\addr \in \addrset$, a JavaScript AST $\tree \in \treeset$, or a function $f
\in \funcset$.  A record $\rec \in \recset$ is a finite mappings from strings to
values.

\subsubsection{Concrete Semantics of Instructions}

\[
  \framebox{$\sem{\inst}: \stset \rightarrow \stset$}
\]

\todo

\subsubsection{Concrete Semantics of Functions}

\[
  \framebox{$\sem{\func}: \stset \times \valset^* \rightarrow \stset$}
\]

\todo

\subsection{Syntatic Views}

\todo

\subsection{Partial Evaluation}

\todo

\subsection{Observational Semantics}

\todo

\subsection{Proof of Semantics Preservation}

\todo
